Parsing Postfix log files

### 2008-10-30 doc/logparser-implementation.tex
Fixed all overfull hbox warnings, and made some small improvements in the process.

        XXX SHOULD THIS LIST BE HERE\@?  MAYBE MOVE TO AFTER THE FIELD
        LIST\@?  DO I NEED THIS LIST AT ALL\@?

@@

            \item [EXPIRED] The mail has been in the queue for five
                days\footnotemark[\value{footnote}]XXX CHECK IF THIS
                FOOTNOTE MAKES SENSE without successful delivery.  A bounce
                mail will be generated and sent to the sender address.

@@

            \item [POSTFIX\_RELOAD] The administrator has instructed
                Postfix to start or stop, and all existing \daemon{smtpd}
                processes will be terminated.  This does not negatively
                affect the log files or mail queued by Postfix for
                delivery.

@@

    \item [result\_data] Sometimes rules need to save a result that is not
        present in the log line: e.g.\ setting \texttt{smtp\_code} when
        mail is accepted.  The format is: \newline{} \tab{} \texttt{XXX
        UPDATE EXAMPLE\@; smtp\_code = 550; recipient = 2, sender = 4;}
        \newline{} i.e.\ semi-colon or comma separated assignment
        statements, with the variable name on the left and the value on the
        right hand side.\footnote{Commas and semi-colons cannot be escaped
        and thus cannot be used.  This feature is intended for use with
        small amounts of data rather than large amounts in any one rule, so
        dealing with escape sequences was deemed unnecessary.}  Any field
        in the results table can be set in this way; additionally
        \texttt{child} and \texttt{pid} are accepted and used respectively
        by the \texttt{TRACK}, \texttt{BOUNCE} and \texttt{SMTPD\_DIED}
        actions, but do not cause data to be entered in the database.

@@

\textbf{Field}      & \textbf{Value}                                    \\
name                & Unknown sender domain                             \\
description         & We do not accept mail from unknown domains        \\
restriction\_name   & reject\_unknown\_sender\_domain                   \\
postfix\_action     & REJECTED                                          \\
program             & \daemon{smtpd}                                    \\
XXX                 & UPDATE REGEX WHEN AUTOMATIC SAVING IS FINISHED    \\
regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                    & \verb!Sender address rejected: Domain not found;! \\
                    & \verb!from=<\5> to=<(__RECIPIENT__)> !            \\
                    & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
result\_data        &                                                   \\
connection\_data    &                                                   \\
action              & DELIVERY\_REJECTED                                \\
hits                & 0                                                 \\
hits\_total         & 0                                                 \\
priority            & 0                                                 \\
%cluster\_group      & 400                                               \\

@@

    \item [Classify words based on their frequency]  The frequency of each
        \texttt{(word, word's position within the log line)} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point)
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that should be noticed
        by the user when reviewing the new \regexes{}.  Variable terms are
        replaced by \texttt{.+}, which means to match zero or more of any
        character.  XXX SHOULD I USE \verb![\S]+! INSTEAD\@?

@@

    \item XXX IS THIS CORRECT NOW\@?  Are there two or more \daemon{smtpd}
        log lines?  Does the second result have a postfix\_action attribute
        of ACCEPTED\@?  The first two \daemon{smtpd} log lines will be a
        connection log line and a mail acceptance log line (Postfix logs
        acceptance as soon as the first recipient has been accepted).

@@

    \item [\daemon{local}, \daemon{pickup}, \daemon{smtp}:] Mail submitted
        locally on the server, \newline{} both local and remote delivery.

>>>>

        XXX MAYBE MOVE THIS LIST TO AFTER THE FIELD LIST\@?  DO I NEED THIS
        LIST AT ALL\@?

@@

            \item [EXPIRED] The mail has been in the queue for five
                days\footnotemark[\value{footnote}] XXX CHECK IF THIS
                FOOTNOTE WORKS without successful delivery.  A bounce mail
                will be generated and sent to the sender address.

@@

            \item [POSTFIX\_RELOAD] The administrator has instructed
                Postfix to start or stop, and all \daemon{smtpd} processes
                will be terminated.  This does not negatively affect the
                log files or mail queued by Postfix for delivery.

@@

    \item [result\_data] Sometimes rules need to save a result that is not
        present in the log line: e.g.\ setting \texttt{smtp\_code} when
        mail is accepted.  The format is: \newline{} \tab{} \texttt{XXX
        UPDATE EXAMPLE\@; smtp\_code = 550; \newline{} recipient = 2,
        sender = 4;} \newline{} i.e.\ semi-colon or comma separated
        assignment statements, with the variable name on the left and the
        value on the right hand side.\footnote{Commas and semi-colons
        cannot be escaped and thus cannot be used.  This feature is
        intended for use with small amounts of data rather than large
        amounts in any one rule, so dealing with escape sequences was
        deemed unnecessary.}  Any field in the results table can be set in
        this way; additionally \texttt{child} and \texttt{pid} are accepted
        and used respectively by the \texttt{TRACK}, \texttt{BOUNCE} and
        \texttt{SMTPD\_DIED} actions, but do not cause data to be entered
        in the database.

@@

\textbf{Field}      & \textbf{Value}                                    \\
name                & Unknown sender domain                             \\
description         & We do not accept mail from unknown domains        \\
restriction\_name   & reject\_unknown\_sender\_domain                   \\
postfix\_action     & REJECTED                                          \\
program             & \daemon{smtpd}                                    \\
XXX                 & UPDATE REGEX FOR AUTOMATIC SAVING                 \\
regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                    & \verb!Sender address rejected: Domain not found;! \\
                    & \verb!from=<\5> to=<(__RECIPIENT__)> !            \\
                    & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
result\_data        &                                                   \\
connection\_data    &                                                   \\
action              & DELIVERY\_REJECTED                                \\
hits                & 0                                                 \\
hits\_total         & 0                                                 \\
priority            & 0                                                 \\
%cluster\_group      & 400                                               \\

@@

    \item [Classify words based on their frequency]  The frequency of every
        tuple \texttt{(word, word's position within the log line)} is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point)
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that should be noticed
        by the user when reviewing the new \regexes{}.  Variable terms are
        replaced by \texttt{.+}, which means to match zero or more of any
        character.  XXX SHOULD I USE \verb![\S]+! INSTEAD\@?

@@

    \item XXX IS THIS CORRECT NOW\@?\newline{}  Are there two or more
        \daemon{smtpd} log lines?  Does the second result have a
        postfix\_action attribute of ACCEPTED\@?  The first two
        \daemon{smtpd} log lines will be a connection log line and a mail
        acceptance log line (Postfix logs acceptance as soon as the first
        recipient has been accepted).

@@

    \item [\daemon{local}, \daemon{pickup}, \daemon{smtp}:] Mail submitted
        locally \newline{} on the server, both local and remote delivery.
